















<!doctype html>
<html lang="en">
  <head>
    <meta charset="utf-8">
    <meta http-equiv="X-UA-Compatible" content="IE=edge">
    <meta http-equiv="last-modified" content="2020-10-09 11:36:08 +0000">
    <meta name="viewport" content="width=device-width, initial-scale=1">
    <!-- meta "search-domain" used for google site search function google_search() -->
    <meta name="search-domain" value="">
    <link rel="stylesheet" type="text/css" href="../assets/css/bootstrap.css" />
    <link rel="stylesheet" type="text/css" href="../assets/css/bootstrap-theme.css" />
    <link rel="stylesheet" type="text/css" href="../assets/css/lesson.css" />
    <link rel="stylesheet" type="text/css" href="../assets/css/syntax.css" />
    <link rel="license" href="#license-info" />

    



    <!-- Favicons for everyone -->
    <link rel="apple-touch-icon-precomposed" sizes="57x57" href="../assets/favicons/dc/apple-touch-icon-57x57.png" />
    <link rel="apple-touch-icon-precomposed" sizes="114x114" href="../assets/favicons/dc/apple-touch-icon-114x114.png" />
    <link rel="apple-touch-icon-precomposed" sizes="72x72" href="../assets/favicons/dc/apple-touch-icon-72x72.png" />
    <link rel="apple-touch-icon-precomposed" sizes="144x144" href="../assets/favicons/dc/apple-touch-icon-144x144.png" />
    <link rel="apple-touch-icon-precomposed" sizes="60x60" href="../assets/favicons/dc/apple-touch-icon-60x60.png" />
    <link rel="apple-touch-icon-precomposed" sizes="120x120" href="../assets/favicons/dc/apple-touch-icon-120x120.png" />
    <link rel="apple-touch-icon-precomposed" sizes="76x76" href="../assets/favicons/dc/apple-touch-icon-76x76.png" />
    <link rel="apple-touch-icon-precomposed" sizes="152x152" href="../assets/favicons/dc/apple-touch-icon-152x152.png" />
    <link rel="icon" type="image/png" href="../assets/favicons/dc/favicon-196x196.png" sizes="196x196" />
    <link rel="icon" type="image/png" href="../assets/favicons/dc/favicon-96x96.png" sizes="96x96" />
    <link rel="icon" type="image/png" href="../assets/favicons/dc/favicon-32x32.png" sizes="32x32" />
    <link rel="icon" type="image/png" href="../assets/favicons/dc/favicon-16x16.png" sizes="16x16" />
    <link rel="icon" type="image/png" href="../assets/favicons/dc/favicon-128.png" sizes="128x128" />
    <meta name="application-name" content="Data Carpentry - R for Social Scientists"/>
    <meta name="msapplication-TileColor" content="#FFFFFF" />
    <meta name="msapplication-TileImage" content="../assets/favicons/dc/mstile-144x144.png" />
    <meta name="msapplication-square70x70logo" content="../assets/favicons/dc/mstile-70x70.png" />
    <meta name="msapplication-square150x150logo" content="../assets/favicons/dc/mstile-150x150.png" />
    <meta name="msapplication-wide310x150logo" content="../assets/favicons/dc/mstile-310x150.png" />
    <meta name="msapplication-square310x310logo" content="../assets/favicons/dc/mstile-310x310.png" />


    <!-- HTML5 shim and Respond.js for IE8 support of HTML5 elements and media queries -->
    <!-- WARNING: Respond.js doesn't work if you view the page via file:// -->
    <!--[if lt IE 9]>
	<script src="https://oss.maxcdn.com/html5shiv/3.7.2/html5shiv.min.js"></script>
	<script src="https://oss.maxcdn.com/respond/1.4.2/respond.min.js"></script>
	<![endif]-->

  <title>
  Code Handout - Starting with Data &ndash; R for Social Scientists
  </title>

  </head>
  <body>

    







    <div class="container">
      
















  
  










<nav class="navbar navbar-default">
  <div class="container-fluid">
    <div class="navbar-header">
      <button type="button" class="navbar-toggle collapsed" data-toggle="collapse" data-target="#bs-example-navbar-collapse-1" aria-expanded="false">
        <span class="sr-only">Toggle navigation</span>
        <span class="icon-bar"></span>
        <span class="icon-bar"></span>
        <span class="icon-bar"></span>
      </button>

      
      
      <a href="http://datacarpentry.org" class="pull-left">
        <img class="navbar-logo" src="../assets/img/dc-icon-black.svg" alt="Data Carpentry logo" />
      </a>
      

      
      <a class="navbar-brand" href="../">Home</a>

    </div>
    <div class="collapse navbar-collapse" id="bs-example-navbar-collapse-1">
      <ul class="nav navbar-nav">

	
        <li><a href="../CODE_OF_CONDUCT.html">Code of Conduct</a></li>

        
	
        <li><a href="../setup.html">Setup</a></li>

        
        
        <li class="dropdown">
          <a href="../" class="dropdown-toggle" data-toggle="dropdown" role="button" aria-haspopup="true" aria-expanded="false">Episodes <span class="caret"></span></a>
          <ul class="dropdown-menu">
            
            
            <li><a href="../00-intro/index.html">Before we Start</a></li>
            
            
            <li><a href="../01-intro-to-r/index.html">Introduction to R</a></li>
            
            
            <li><a href="../02-starting-with-data/index.html">Starting with Data</a></li>
            
            
            <li><a href="../03-dplyr-tidyr/index.html">Data Wrangling with dplyr and tidyr</a></li>
            
            
            <li><a href="../04-ggplot2/index.html">Data Visualisation with ggplot2</a></li>
            
            
            <li><a href="../0x-json/index.html">Processing JSON data</a></li>
            
	    <li role="separator" class="divider"></li>
            <li><a href="../aio/index.html">All in one page (Beta)</a></li>
          </ul>
        </li>
        
	

	
	
        <li class="dropdown">
          <a href="../" class="dropdown-toggle" data-toggle="dropdown" role="button" aria-haspopup="true" aria-expanded="false">Extras <span class="caret"></span></a>
          <ul class="dropdown-menu">
            <li><a href="../reference.html">Reference</a></li>
            
            
            <li><a href="../about/index.html">About</a></li>
            
            
            <li><a href="../data-visualisation-handout/index.html">Code Handout - Data Visualisation with ggplot2</a></li>
            
            
            <li><a href="../data-wrangling-handout/index.html">Code Handout</a></li>
            
            
            <li><a href="../discuss/index.html">Discussion</a></li>
            
            
            <li><a href="../figures/index.html">Figures</a></li>
            
            
            <li><a href="../guide/index.html">Instructor Notes</a></li>
            
            
            <li><a href="../intro-R-handout/index.html">Code Handout - Introduction to R</a></li>
            
            
            <li><a href="../starting-with-data-handout/index.html">Code Handout - Starting with Data</a></li>
            
          </ul>
        </li>
	

	
        <li><a href="../LICENSE.html">License</a></li>
	
	<li><a href="/edit//_extras/starting-with-data-handout.Rmd" data-checker-ignore>Improve this page <span class="glyphicon glyphicon-pencil" aria-hidden="true"></span></a></li>
	
      </ul>
      <form class="navbar-form navbar-right" role="search" id="search" onsubmit="google_search(); return false;">
        <div class="form-group">
          <input type="text" id="google-search" placeholder="Search..." aria-label="Google site search">
        </div>
      </form>
    </div>
  </div>
</nav>
















<h1 class="maintitle"><a href="../">R for Social Scientists</a>: Code Handout - Starting with Data</h1>



<article>
```{r, include = FALSE}
knitr::opts_chunk$set(fig.width = 3, 
                  fig.height = 3, 
                  message = FALSE, 
                  warning = FALSE,
                  eval = FALSE)

```

This document contains all of the functions that were covered in the 
_Introduction to R_ workshop. Each function is presented alongside an example of
how it can be used. 

All of the examples below are in the context of the Palmer Penguins, found 
[here (link)](https://allisonhorst.github.io/palmerpenguins/index.html). 


## Packages

- `library()` -- loads packages into your `R` session
```{r, message=FALSE, warning=FALSE}
library(tidyverse)
library(lubridate)
```


## Importing Data
- `read_csv()` -- function to import a csv file. 
  * First argument is the path to the data, passed as a character 
  (inside quotations). 
  * You can specify what values should be considered missing, using the `na`
  argument. 
```{r}
penguins <- read_csv("data/penguins.csv")
```


## Inspecting Data

- `dim()` - returns a vector with the number of rows as the first element,
and the number of columns as the second element (the **dim**ensions of
the object)
```{r}
dim(penguins)
```

- `nrow()` - returns the number of rows
- `ncol()` - returns the number of columns
```{r}
nrow(penguins)
ncol(penguins)
```

- `head()` - displays the first 6 rows of the dataframe
- `tail()` - displays the last 6 rows of the dataframe
```{r}
head(penguins)
tail(penguins)
```

- `names()` - returns the all of the names of an object (both row and column) 
- `colnames()` - returns column names for dataframes (without row names)
```{r}
names(penguins)
colnames(penguins)
```

- `glimpse()` - provides a preview of the data, where column names are presented 
with their associated data types, and the entries from each column are printed 
in each row 
```{r}
glimpse(penguins)
```

- `str()` - returns the structure of the object and information about the class,
the names and data types of each column, and a preview of the first entries of
each column
```{r}
str(penguins)
```

- `summary()` - provides summary statistics for each column
  * Note: summary statistics for character variables are not meaningful, as they 
  simply state the number of observations (length) of the variable
```{r}
summary(penguins)
```


## Subsetting Data

- `[]` -- selects rows and columns from a dataframe
  * The first entry is the row number, the second entry is the column number(s),
  and they are separated with a comma. 
```{r}
## Selects the element in the first row, second column
penguins[1, 2]

## Selects every element in the fourth row
penguins[4, ]

## Selects every element in the third column
penguins[, 3]
```

- `[[]]` -- selects a column from a dataframe
  * Inside the brackets you can pass either the number of the column or the 
  name of the column (in quotations)
```{r}
penguins[[1]]

penguins[["island"]]
```

- `$` -- selects a column from a dataframe, where the name of the dataframe is 
on the left and the name of the column is on the right
```{r}
penguins$body_mass_g
```


## Working with Different Data Types

- `factor()` -- creates a categorical variable from a character or numeric 
variable, variable has a factor datatype
  * the values (level) of the factor levels is specified in the `levels` 
  argument, where the levels must be specified in a vector (using `c()`)
  * Note: the order you wish for the levels to appear is how you should list 
  them in the `levels` argument, you can also specify `ordered = TRUE` to 
  ensure the levels remain in this order
```{r}
penguins$year_fct <- factor(penguins$year, 
                               levels = c("2007", "2008", "2009"), 
                            ordered = TRUE)
```

- `as.factor()` -- creates a categorical variable from a character or numeric 
variable, variable has a factor datatype
  * does not allow for you to specify the order of the levels
  * defaults to alphabetical ordering for factor levels
```{r}
penguins$year_fct <- as.factor(penguins$year)
```

- `levels()` -- returns the levels of a variable with a factor datatype, in the 
order they were stored
  * Note: this function will not work for character datatypes
```{r}
levels(penguins$year_fct)
```

- `nlevels()` -- returns the number of levels of a variable with a factor
datatype
  * Note: this function will not work for character datatypes
```{r}
nlevels(penguins$year_fct)
```

- `as.character()` -- creates a character variable from a numeric or factor 
variable
```{r}
penguins$species_chr <- as.character(penguins$species)
```

- `ymd()` -- transforms dates stored as character or numeric variables to dates
  * Note: to use this function, dates must be stored in year-month-day format
  * The function does well with heterogeneous formats (as seen below), but
  formats where some of the entries are not in double digits may not be parsed 
  correctly.
```{r}
x <- c("2009-01-01", "2009-01-02", "2009-01-03")
ymd(x)
```

- `day()` -- extracts the day (number) of a date variable 
```{r}
day(x)
```

- `month()` -- extracts the month (number) of a date variable
```{r}
month(x)
```

- `year()` -- extracts the year of a date variable
```{r}
year(x)
```


## Visualizing Data

- `plot()` -- a generic function for plotting R objects
  * In this lesson `plot()` was used to create bargraphs of categorical
  variables.
```{r}
plot(penguins$species)
```



</article>

      
      






<footer>
  <div class="row">
    <div class="col-md-6 license" id="license-info" align="left">
	
	Licensed under <a href="">CC-BY 4.0</a> 2018–2020
	by <a href="https://carpentries.org/">The Carpentries</a>
        <br>
        Licensed under <a href="">CC-BY 4.0</a> 2016–2018
	by <a href="http://datacarpentry.org">Data Carpentry</a>
	
    </div>
    <div class="col-md-6 help-links" align="right">
	
	<a href="/edit//_extras/starting-with-data-handout.Rmd" data-checker-ignore>Edit on GitHub</a>
	
	/
	<a href="/blob//CONTRIBUTING.md" data-checker-ignore>Contributing</a>
	/
	<a href="/">Source</a>
	/
	<a href="/blob//CITATION" data-checker-ignore>Cite</a>
	/
	<a href="mailto:team@carpentries.org">Contact</a>
    </div>
  </div>
  <div class="row">
    <div class="col-md-12" align="center">
      Using <a href="https://github.com/carpentries/styles/">The Carpentries style</a>
      version <a href="https://github.com/carpentries/styles/releases/tag/v9.5.3">9.5.3</a>.
    </div>
  </div>
</footer>

      
    </div>
    
<script src="../assets/js/jquery.min.js"></script>
<script src="../assets/js/bootstrap.min.js"></script>
<script src="../assets/js/lesson.js"></script>
<script>
  (function(i,s,o,g,r,a,m){i['GoogleAnalyticsObject']=r;i[r]=i[r]||function(){
  (i[r].q=i[r].q||[]).push(arguments)},i[r].l=1*new Date();a=s.createElement(o),
  m=s.getElementsByTagName(o)[0];a.async=1;a.src=g;m.parentNode.insertBefore(a,m)
  })(window,document,'script','https://www.google-analytics.com/analytics.js','ga');
  ga('create', 'UA-37305346-2', 'auto');
  ga('send', 'pageview');
</script>

  </body>
</html>
