















<!doctype html>
<html lang="en">
  <head>
    <meta charset="utf-8">
    <meta http-equiv="X-UA-Compatible" content="IE=edge">
    <meta http-equiv="last-modified" content="2020-10-09 11:36:08 +0000">
    <meta name="viewport" content="width=device-width, initial-scale=1">
    <!-- meta "search-domain" used for google site search function google_search() -->
    <meta name="search-domain" value="">
    <link rel="stylesheet" type="text/css" href="../assets/css/bootstrap.css" />
    <link rel="stylesheet" type="text/css" href="../assets/css/bootstrap-theme.css" />
    <link rel="stylesheet" type="text/css" href="../assets/css/lesson.css" />
    <link rel="stylesheet" type="text/css" href="../assets/css/syntax.css" />
    <link rel="license" href="#license-info" />

    



    <!-- Favicons for everyone -->
    <link rel="apple-touch-icon-precomposed" sizes="57x57" href="../assets/favicons/dc/apple-touch-icon-57x57.png" />
    <link rel="apple-touch-icon-precomposed" sizes="114x114" href="../assets/favicons/dc/apple-touch-icon-114x114.png" />
    <link rel="apple-touch-icon-precomposed" sizes="72x72" href="../assets/favicons/dc/apple-touch-icon-72x72.png" />
    <link rel="apple-touch-icon-precomposed" sizes="144x144" href="../assets/favicons/dc/apple-touch-icon-144x144.png" />
    <link rel="apple-touch-icon-precomposed" sizes="60x60" href="../assets/favicons/dc/apple-touch-icon-60x60.png" />
    <link rel="apple-touch-icon-precomposed" sizes="120x120" href="../assets/favicons/dc/apple-touch-icon-120x120.png" />
    <link rel="apple-touch-icon-precomposed" sizes="76x76" href="../assets/favicons/dc/apple-touch-icon-76x76.png" />
    <link rel="apple-touch-icon-precomposed" sizes="152x152" href="../assets/favicons/dc/apple-touch-icon-152x152.png" />
    <link rel="icon" type="image/png" href="../assets/favicons/dc/favicon-196x196.png" sizes="196x196" />
    <link rel="icon" type="image/png" href="../assets/favicons/dc/favicon-96x96.png" sizes="96x96" />
    <link rel="icon" type="image/png" href="../assets/favicons/dc/favicon-32x32.png" sizes="32x32" />
    <link rel="icon" type="image/png" href="../assets/favicons/dc/favicon-16x16.png" sizes="16x16" />
    <link rel="icon" type="image/png" href="../assets/favicons/dc/favicon-128.png" sizes="128x128" />
    <meta name="application-name" content="Data Carpentry - R for Social Scientists"/>
    <meta name="msapplication-TileColor" content="#FFFFFF" />
    <meta name="msapplication-TileImage" content="../assets/favicons/dc/mstile-144x144.png" />
    <meta name="msapplication-square70x70logo" content="../assets/favicons/dc/mstile-70x70.png" />
    <meta name="msapplication-square150x150logo" content="../assets/favicons/dc/mstile-150x150.png" />
    <meta name="msapplication-wide310x150logo" content="../assets/favicons/dc/mstile-310x150.png" />
    <meta name="msapplication-square310x310logo" content="../assets/favicons/dc/mstile-310x310.png" />


    <!-- HTML5 shim and Respond.js for IE8 support of HTML5 elements and media queries -->
    <!-- WARNING: Respond.js doesn't work if you view the page via file:// -->
    <!--[if lt IE 9]>
	<script src="https://oss.maxcdn.com/html5shiv/3.7.2/html5shiv.min.js"></script>
	<script src="https://oss.maxcdn.com/respond/1.4.2/respond.min.js"></script>
	<![endif]-->

  <title>
  Code Handout - Introduction to R &ndash; R for Social Scientists
  </title>

  </head>
  <body>

    







    <div class="container">
      
















  
  










<nav class="navbar navbar-default">
  <div class="container-fluid">
    <div class="navbar-header">
      <button type="button" class="navbar-toggle collapsed" data-toggle="collapse" data-target="#bs-example-navbar-collapse-1" aria-expanded="false">
        <span class="sr-only">Toggle navigation</span>
        <span class="icon-bar"></span>
        <span class="icon-bar"></span>
        <span class="icon-bar"></span>
      </button>

      
      
      <a href="http://datacarpentry.org" class="pull-left">
        <img class="navbar-logo" src="../assets/img/dc-icon-black.svg" alt="Data Carpentry logo" />
      </a>
      

      
      <a class="navbar-brand" href="../">Home</a>

    </div>
    <div class="collapse navbar-collapse" id="bs-example-navbar-collapse-1">
      <ul class="nav navbar-nav">

	
        <li><a href="../CODE_OF_CONDUCT.html">Code of Conduct</a></li>

        
	
        <li><a href="../setup.html">Setup</a></li>

        
        
        <li class="dropdown">
          <a href="../" class="dropdown-toggle" data-toggle="dropdown" role="button" aria-haspopup="true" aria-expanded="false">Episodes <span class="caret"></span></a>
          <ul class="dropdown-menu">
            
            
            <li><a href="../00-intro/index.html">Before we Start</a></li>
            
            
            <li><a href="../01-intro-to-r/index.html">Introduction to R</a></li>
            
            
            <li><a href="../02-starting-with-data/index.html">Starting with Data</a></li>
            
            
            <li><a href="../03-dplyr-tidyr/index.html">Data Wrangling with dplyr and tidyr</a></li>
            
            
            <li><a href="../04-ggplot2/index.html">Data Visualisation with ggplot2</a></li>
            
            
            <li><a href="../0x-json/index.html">Processing JSON data</a></li>
            
	    <li role="separator" class="divider"></li>
            <li><a href="../aio/index.html">All in one page (Beta)</a></li>
          </ul>
        </li>
        
	

	
	
        <li class="dropdown">
          <a href="../" class="dropdown-toggle" data-toggle="dropdown" role="button" aria-haspopup="true" aria-expanded="false">Extras <span class="caret"></span></a>
          <ul class="dropdown-menu">
            <li><a href="../reference.html">Reference</a></li>
            
            
            <li><a href="../about/index.html">About</a></li>
            
            
            <li><a href="../data-visualisation-handout/index.html">Code Handout - Data Visualisation with ggplot2</a></li>
            
            
            <li><a href="../data-wrangling-handout/index.html">Code Handout</a></li>
            
            
            <li><a href="../discuss/index.html">Discussion</a></li>
            
            
            <li><a href="../figures/index.html">Figures</a></li>
            
            
            <li><a href="../guide/index.html">Instructor Notes</a></li>
            
            
            <li><a href="../intro-R-handout/index.html">Code Handout - Introduction to R</a></li>
            
            
            <li><a href="../starting-with-data-handout/index.html">Code Handout - Starting with Data</a></li>
            
          </ul>
        </li>
	

	
        <li><a href="../LICENSE.html">License</a></li>
	
	<li><a href="/edit//_extras/intro-R-handout.Rmd" data-checker-ignore>Improve this page <span class="glyphicon glyphicon-pencil" aria-hidden="true"></span></a></li>
	
      </ul>
      <form class="navbar-form navbar-right" role="search" id="search" onsubmit="google_search(); return false;">
        <div class="form-group">
          <input type="text" id="google-search" placeholder="Search..." aria-label="Google site search">
        </div>
      </form>
    </div>
  </div>
</nav>
















<h1 class="maintitle"><a href="../">R for Social Scientists</a>: Code Handout - Introduction to R</h1>



<article>
```{r, include = FALSE}
knitr::opts_chunk$set(fig.width = 3, 
                  fig.height = 3, 
                  message = FALSE, 
                  warning = FALSE,
                  eval = FALSE)

```

This document contains all of the functions that were covered in the 
_Introduction to R_ workshop. Each function is presented alongside an example of
how it can be used. 


## Creating Objects 

- `<-` -- "assignment arrow", assigns a value (vector, dataframe, single value)
to the name of a variable 
```{r}

x <- 3
y <- c(1, 2, 3)
z <- x + y
```

- `c()` -- the "concatenate" function combines inputs to form a vector, the 
values have to be the same data type. 
```{r}
animals <- c("bird", "cat", "dog")
numbers <- c(1, 14, 57, 89)
logicals <- c(TRUE, FALSE, TRUE, TRUE)
```


## Inspecting Objects

- `str()` -- compact display of the structure of an R object 
```{r}
str(animals)
```

- `class()` -- returns the type of element of any R object
```{r}
class(logicals)
```

- `typeof()` -- returns the data type or storage mode of any R object
```{r}
typeof(numbers)
```


## Functions in R 

- `args()` -- returns the arguments of a function
```{r}
args(round)
```

- named arguments -- the name of the argument the function expects
  * You can choose to not name your arguments, __if__ you know the __exact__ 
  order they should be in! 
  * However, we generally discourage this.
```{r}
## Either of these work, since the digits argument is named explicitly.
round(3.14159, digits = 2)
round(digits = 2, 3.14159)

## This does not work, since the arguments are not named and in the incorrect order. 
round(2, 3.14159)
```


## Functions to Summarize Data

- `sqrt()` -- returns the square root of a numeric variable
```{r}
sqrt(numbers)
```

- `mean()` -- returns the mean of a numeric variable
  * You can add the `na.rm` argument, to remove `NA` values before calculating 
  the mean.
```{r}
sqrt(numbers)
```

- `max()` -- returns the maximum of a numeric variable
  * You can add the `na.rm` argument, to remove `NA` values before calculating 
  the max.
```{r}
sqrt(numbers)
```

- `sum()` -- returns the sum of a numeric variable
  * You can add the `na.rm` argument, to remove `NA` values before calculating 
  the sum.
```{r}
sqrt(numbers)
```

- `length()` -- returns the length of a vector (of any datatype)
```{r}
length(animals)
```


## Subsetting Data

- `[]` -- used to subset elements from a vector
```{r}
animals[3]
## selects the third element

animals[2:3]
## selects the second and third element

animals[c(1, 3)]
## selects the first and third element
```

- relational operators -- return logical values indicating where a relation is 
satisfied. The most commonly used logical operators for data analysis are as follows:
  * `==` means "equal to"
  * `!=` means "not equal to"
  * `>` or `<` means "greater than" or "less than"
  * `>=` or `<=` means "greater than or equal to" or "less than or equal to"
```{r}
animals == "dog"

animals != "cat"

numbers > 4

numbers <= 12
```


- logical operators -- join subset criteria together 
  * `&` means "and" -- where two criteria must **both** be satisfied
  * `|` means "or" -- where at least one criteria must be satisfied
```{r}
numbers > 4 & numbers < 20

animals == "dog" | animals == "cat"
```
  

- `%in%` -- the "inclusion operator", allows you to test if any of the elements
of a search vector (on the left hand side) are found in the target vector (on
the right hand side). 
  * The levels of the target vector must be included in a vector (`c()`). 
```{r}
possessions <- c("car", "bicycle", "radio", "television", "mobile_phone")

possessions %in% c("car", "bicycle", "motorcycle")
```


## Missing Data 

- `is.na()` -- returns a vector of logical values indicating which elements of 
a vector have `NA` values
  * Often combined with `!`, where the `!` negates the previous statement (e.g. 
  `!TRUE` is equal to `FALSE`). 
```{r}
missing <- c(1, 3, NA, 7, 12, NA)

is.na(missing)

!is.na(missing)
```

- `na.omit()` -- removes the observations with `NA` values
```{r}
na.omit(missing)
```

- `complete.cases()` -- returns a vector of logical values indicating which
elements of a vector **are not** missing (`NA`) values
```{r}
complete.cases(missing)
```


</article>

      
      






<footer>
  <div class="row">
    <div class="col-md-6 license" id="license-info" align="left">
	
	Licensed under <a href="">CC-BY 4.0</a> 2018–2020
	by <a href="https://carpentries.org/">The Carpentries</a>
        <br>
        Licensed under <a href="">CC-BY 4.0</a> 2016–2018
	by <a href="http://datacarpentry.org">Data Carpentry</a>
	
    </div>
    <div class="col-md-6 help-links" align="right">
	
	<a href="/edit//_extras/intro-R-handout.Rmd" data-checker-ignore>Edit on GitHub</a>
	
	/
	<a href="/blob//CONTRIBUTING.md" data-checker-ignore>Contributing</a>
	/
	<a href="/">Source</a>
	/
	<a href="/blob//CITATION" data-checker-ignore>Cite</a>
	/
	<a href="mailto:team@carpentries.org">Contact</a>
    </div>
  </div>
  <div class="row">
    <div class="col-md-12" align="center">
      Using <a href="https://github.com/carpentries/styles/">The Carpentries style</a>
      version <a href="https://github.com/carpentries/styles/releases/tag/v9.5.3">9.5.3</a>.
    </div>
  </div>
</footer>

      
    </div>
    
<script src="../assets/js/jquery.min.js"></script>
<script src="../assets/js/bootstrap.min.js"></script>
<script src="../assets/js/lesson.js"></script>
<script>
  (function(i,s,o,g,r,a,m){i['GoogleAnalyticsObject']=r;i[r]=i[r]||function(){
  (i[r].q=i[r].q||[]).push(arguments)},i[r].l=1*new Date();a=s.createElement(o),
  m=s.getElementsByTagName(o)[0];a.async=1;a.src=g;m.parentNode.insertBefore(a,m)
  })(window,document,'script','https://www.google-analytics.com/analytics.js','ga');
  ga('create', 'UA-37305346-2', 'auto');
  ga('send', 'pageview');
</script>

  </body>
</html>
